Solar wind predictions for the Parker Solar Probe orbit

### 2017-10-23 paperMVVB/sections/acknowledgments.tex
referee and w/o figures versions added
adopted referee comments; made changes bold face
#2 integrated changes by MVVB

\begin{acknowledgements}
	The authors acknowledge support of the Coronagraphic German~and~US SolarProbePlus Survey (CGAUSS) project for WISPR by the German Aerospace Center (DLR) under grant 50~OL~1601 as national contribution to the Parker Solar Probe mission. The authors thank the Helios and OMNI PIs/teams for creating and making available the solar wind in situ data. The Helios and the OMNI data are supplied by the NASA Space Science Data Coordinated Archive (NSSDCA) and the Space Physics Data Facility (SPDF) at NASA's Goddard Space Flight Center (GSFC). Additional thanks for maintaining and providing the international sunspot number series goes to the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO) at the Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB). The PSP SPICE kernel was kindly provided by Angelos~Vourlidas.
%	We thank the referees for /very important comments and suggestions. /careful review of this paper.

%	facilities/persons which made available the Helios and OMNI data\\
% 	``Ian Richardson and Hilary Cane for making their Interplanetary Coronal Mass Ejection list readily available''\\
% 	``Joseph King and Natalia Papitashvili for creating the combined OMNI data set''\\
% 	``We thank the reviewers for the careful review of this paper''\\
% 	The Authors extend their thanks to the referee for important comments and suggestions.\\
% 	We would like to thank XX for reading drafts of this manuscript...\\
% 	We would like to thank the Helios team and NASA's SPDF for supplying the plasma and magnetic field data (url)\\
\end{acknowledgements}

>>>>

\begin{acknowledgements}
	The authors acknowledge support of the Coronagraphic German and US SolarProbePlus Survey (CGAUSS) project for WISPR by the German Aerospace Center (DLR) under grant 50~OL~1601 as national contribution to the Parker Solar Probe mission. The authors thank the Helios and OMNI PIs/teams for creating and making available the solar wind in situ data. The Helios and the OMNI data are supplied by the NASA Space Science Data Coordinated Archive (NSSDCA) and the Space Physics Data Facility (SPDF) at NASA's Goddard Space Flight Center (GSFC). Additional thanks for maintaining and providing the international sunspot number series goes to the World Data Center -- Sunspot Index and Long-term Solar Observations (WDC-SILSO) at the Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium (ROB). The PSP SPICE kernel was kindly provided by Angelos~Vourlidas. \textbf{The authors thank the referee for the careful review of this manuscript and helpful comments and suggestions.}
%	We thank the referees for /very important comments and suggestions. /careful review of this paper.
%	facilities/persons which made available the Helios and OMNI data\\
% 	``Ian Richardson and Hilary Cane for making their Interplanetary Coronal Mass Ejection list readily available''\\
% 	``Joseph King and Natalia Papitashvili for creating the combined OMNI data set''\\
% 	``We thank the reviewers for the careful review of this paper''\\
% 	The Authors extend their thanks to the referee for important comments and suggestions.\\
% 	We would like to thank XX for reading drafts of this manuscript...\\
% 	We would like to thank the Helios team and NASA's SPDF for supplying the plasma and magnetic field data (url)\\
\end{acknowledgements}

### 2017-10-23 paperMVVB/sections/content.tex
#2 integrated changes by MVVB

	\caption{Fit coefficients \textbf{for the distance depending single lognormal function, equation (\ref{eq:single_lognormal_fit_function}) combined with (\ref{eq:power_function})}, respectively double lognormal \textbf{function (\ref{eq:double_lognormal_fit_function})} for the velocity from the combined Helios data. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value.} The seasonal variations are calculated from Earth's orbital solar distance variation and the derived exponents.}

@@

\textbf{The power law lognormal models and the power law double lognormal model for the velocity, which result from the fitting, are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot} together with their median values.} The model’s magnetic field strength is broader around values of \SI{40}{nT} at the lower distance boundary than the data's frequency distribution implies. This behavior is expected because of the distance independent shape approximation applied. The velocity and temperature models’ upper values generally show a higher abundance than the actual data, see also zoom boxes in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}. \textbf{The increasing high velocity tail with distance} comes from using the same exponent for both slow and fast components. This effect is not seen in the data, more specifically, not only the slowest wind but also the fastest wind is expected to converge to more average speeds \citep{Sanchez-Diaz2016}.
%\citep{Sanchez-Diaz2016} p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction)

@@

%\subsection{Distance scaling law variations}
%\label{sec:distance_scaling_law_variations}
\textbf{This solar wind model} for the radial distance dependence is representative for the time of the Helios observations around the rise of solar cycle~21. The variation of yearly power law fit exponents \textbf{by fitting the solar distance dependency (\ref{eq:power_function})} are shown in Fig.~\ref{fig:yearly_gradients_c} together with the yearly SSN for the time period \numrange{1974}{1982}. It can be seen that during the Helios time period there might be some systematic  variation of the exponents with solar activity---at least for the velocity and temperature exponents.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/yearly_gradients_c.pdf}}
	\caption{Helios yearly variation of the solar wind \textbf{parameter power exponents for the dependence on radial distance} together with the SIDC 13-month smoothed monthly SSN. The weighted standard deviations and average values for all years are indicated by the shaded areas. In this plot the 21~days since Helios launch in the year 1974 are omitted because a distance range of merely \SIrange{0.95}{0.98}{\au} was covered that year.}
	\label{fig:yearly_gradients_c}
\end{figure}
%yearly variation of the solar wind parameters' average fit function exponents
However, for simplicity we assume, that the distance scaling laws can be treated as time independent and include the calculated exponents’ yearly variations $\Delta e$, summarized in Table~\ref{tab:mean_median_fit_parameter}, as relative uncertainties.

Since we neglect possible variations of the distance scaling laws, we combine the frequency distribution’s median solar activity dependency (\ref{eq:median_with_ssn}) derived for \SI{1}{\au} from the OMNI data with the power law exponents (\ref{eq:power_function}) derived from the Helios data:
\begin{align}
	x_\text{med}(ssn,r) &= (a_\text{med} \cdot ssn + b_\text{med}) \cdot r^{e'}	\,.	\label{eq:general_sw_model}
\end{align}
Thus\textbf{, implementing the median and average relations into the lognormal function (\ref{eq:single_lognormal_fit_function}),} we obtain the combined model function $W'''(x,ssn,r)$ and for the velocity $W_\text{II}'''(x,ssn,r)$ with the double lognormal function (\ref{eq:double_lognormal_fit_function}). \textbf{The corresponding median and average relations for each solar wind parameter, with the values resulting from our analyses, are listed in the following. Their numerical values are the fit parameters from Table~\ref{tab:ssn_fit_parameters} and the exponents from Table~\ref{tab:extrapolation_model_fit_parameters}.}

\textbf{\begin{itemize}
	\item The magnetic field strength relations, depending on solar activity and solar distance, are:
	\begin{align}
		B_\text{med}(ssn,r) &= (\SI{0.0131}{\nT} \cdot ssn + \SI{4.29}{\nT}) \cdot r^{-1.66}	\,,	\label{eq:median_B}\\
		B_\text{avg}(ssn,r) &= 1.0879 \cdot B_\text{med}(ssn,r)\,.
	\end{align}
	\item The proton velocity relations for the slow and fast components, depending on solar distance, are:
	\begin{align}
		v_\text{med}^\text{slow}(r) &= \SI{363}{\km\per\s} \cdot r^{0.099}	\,,&	&	&v_\text{med}^\text{fast}(r) &= \SI{483}{\km\per\s} \cdot r^{0.099}	\,,	\label{eq:median_v}\\
		v_\text{avg}^\text{slow}(r) &= 1.0101 \cdot v_\text{med}^\text{slow}(r)	\,,&	&	&v_\text{avg}^\text{fast}(r) &= 1.023 \cdot v_\text{med}^\text{fast}(r)	\,.
	\end{align}
	The share of both components is balanced with solar activity:
	\begin{align}
		c(ssn) = -0.00180 \cdot ssn + 0.64	\,.	\label{eq:balance_v}
	\end{align}
	\item The relations of the proton density are:
	\begin{align}
		n_\text{med}(ssn,r) &= \left(\SI{0.0038}{\per\cm\cubed} \cdot ssn + \SI{4.50}{\per\cm\cubed}\right) \cdot r^{-2.11}	\,,	\label{eq:median_n}\\
		n_\text{avg}(ssn,r) &= 1.305 \cdot n_\text{med}(ssn,r)\,.
	\end{align}
	\item The proton temperature relations are:
	\begin{align}
		T_\text{med}(ssn,r) &= (\SI{197}{\K} \cdot ssn + \SI{57300}{\K}) \cdot r^{-1.10}	\,,	\label{eq:median_T}\\
		T_\text{avg}(ssn,r) &= 1.654 \cdot T_\text{med}(ssn,r)\,.
	\end{align}
\end{itemize}}

\textbf{These models average} over seasonal variations because \textbf{they are} based on yearly data. As the OMNI data are time-shifted to the nose of the Earth’s bow shock, this leads to yearly solar distance variations of \SI{+-1.67}{\%} as it orbits the Sun. The resulting maximal solar wind parameter variation amplitudes over the year can thus be derived from the derived power law exponents. They are estimated to be smaller than \SI{4}{\percent} as seen in Tab.~\ref{tab:extrapolation_model_fit_parameters}. \citet{Bruno1986} \textbf{and} \citet{Balogh1999} have pointed out, that the solar wind parameters vary with latitudinal separation from the heliospheric current sheet (HCS). Its position in heliographic latitude is highly variable around the solar equator \citep{Schwenn1990}, furthermore, the Earth’s orbit varies over the course of the year by \SI{+-7.2}{\degree} in latitude. \textbf{Since this latitudinal separation is highly variable, and requires significant effort to calculate for an extended time series, we have ignored this aspect in this analysis.}	%Balogh et al. (p.~162, 1999);	 (Schwenn 1990, p.~126)

@@

	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. \textbf{The models obtained from Helios and OMNI measurements} are extrapolated to the PSP region---for the range between the extreme cases of solar minimum and maximum, i.e., $0 \le ssn \le 200$\textbf{. The lower edges of the shaded areas correspond to solar minimum, the upper edges to maximum.} Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \textbf{\citet{Sheeley1997} and \citet{Wang2000}}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}

@@

The magnetic field strength is found to increase from median values of about \SI{43}{\nT} at \SI{0.25}{\au} to \SI{715}{\nT} at \SI{0.046}{\au} for a SSN of 0. Taking a SSN of 200 increases the values to \SI{69}{\nT} and \SI{1152}{\nT}. Our extrapolation results are slightly flatter than those derived from the analytical magnetic field model by \citet{Banaszkiewicz1998}, who constructed an analytic dipole plus quadrupole plus current sheet (DQCS) model for solar minimum.
\textbf{We cannot compare with their absolute values, because their DQCS model assumes fields of \SI{3.1}{\nT} at \SI{1}{\au} coming from coronal holes with boundaries at \SI{+-60}{\degree} latitudes (it excludes the fields from the equatorial slow wind belt).}

@@

The proton density increases from about \SI{84}{\per\cm\cubed} at \SI{0.25}{\au} to about \SI{3018}{\per\cm\cubed} at \SI{0.046}{\au} \textbf{6~years after a SSN of 0 occurred}. Being almost independent of the SSN the values for a SSN of 200 are only \SI{17}{\%} larger. The results are in good agreement with those of \citet{Leblanc1998}, who derived an electron density model from type~III radio burst observations. Their model shows that the density distance dependency scales with $r^{-2}$ and steepens just below \SI{10}{\Rs} with $r^{-6}$. \textbf{We assumed a solar wind helium abundance of \SI{5}{\%} to convert their electron densities to proton densities.}

@@

%\subsection{PSP solar wind environment estimation}
Implementing the \textbf{SSN predictions} for the PSP mission time and the orbital trajectory data, \textbf{we can infer} which solar wind parameter magnitudes can be expected. Figs.~\ref{fig:SPP_perihelia_prediction_f_plot} and \ref{fig:SPP_perihelia_prediction_nearest_f_plot} show \textbf{the median values (\ref{eq:median_with_ssn}) of} the considered different solar wind parameters for 12-day periods, comprising the first perihelion in Novemver 2018 and the \textbf{first} closest perihelion in December 2024.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) during 12~days in 2018 with PSP's first perihelion at about \SI{0.16}{\au}. For the velocity the combined median is calculated and also the SSN independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_f_plot}
\end{figure}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) \textbf{during} 12~days in 2024 with PSP's \textbf{22nd (the first closest)} perihelion at \SI{0.0459}{\au}. For the velocity the combined median is calculated and also the SSN independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_nearest_f_plot}
\end{figure}
In the beginning of the mission median values of about \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K} are estimated to be measured at \SI{0.16}{\au}, increasing to about \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K} during the \textbf{first} closest approach at \SI{0.046}{\au}. \textbf{Monthly SSNs---shifted by the time lags specific to the solar wind parameters---are used in the calculation of the solar wind predictions. These SSNs are either actual smoothed values from the SIDC with their reported standard deviations, short-term predictions from the SWPC with their expected ranges or actual smoothed values from the SIDC shifted by 11~years with half/twice their values as uncertainties. The error bands given in both figures, calculated by error propagation, include these SSN ranges and the derived fit parameter errors.}
%B
% 87.3642361846083
%N
% 213.613795298952	first
% 2950.67837375615	nearest
% 2925.63248639046	SSN 0
% 3421.59065204775	SSN 200
%T
% 503136.759398494

\textbf{Now the estimated solar wind environment can finally be derived from the function $W'''(x,ssn,r)$. The estimated frequency distributions of the four solar wind parameters at PSP's first and 22nd (first closest) perihelia are plotted in Fig.~\ref{fig:SPP_sw_distributions}.}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_sw_distributions.pdf}}
	\caption{\textbf{Frequency distributions of the four solar wind parameters (same as in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}) and those estimated with the solar wind model for PSP's first and 22nd (first closest) perihelia. In these figures the frequencies of both extrapolated curves are scaled for visibility to the same height as the \SI{1}{\au} distribution.}}
	\label{fig:SPP_sw_distributions}
\end{figure}
\textbf{Note again that the velocity and temperature distributions for the 22nd perihelion are only upper limits and the actual by PSP to be encountered values are expected to be smaller.}

\section{Discussion}
\label{sec:discussion}
\textbf{The scientific objective of this study, being part of the German contribution Coronagraphic German and US Solar Probe Plus Survey (CGAUSS) to the Wide field Imager for Solar Probe (WISPR), is to model the solar wind environment for the Parker Solar Probe mission (formerly Solar probe Plus) to be launched mid 2018.}

@@

\textbf{For determining solar activity and solar distance dependent relations exclusively for the median and average solar wind values, we would have used the simpler approach of combining the radial dependence of averaged Helios data with averaged \SI{1}{\au} OMNI data---scaled with the SSN. The results of this simpler analysis method would be similar (at least the distance scaling results are similar; compare the exponents in Tables~\ref{tab:mean_median_fit_parameter} and \ref{tab:extrapolation_model_fit_parameters}) and would be sufficient for this purpose. Though, we are not only interested in the averages but in the bulk distributions, that is, the whole range of values which might occur. For the determination of the frequency distributions the use of the more complex fit model is important, because the distance between median and average values determines the width of the lognormal distributions. With our approach we are able to control and fit this distance.}

\textbf{It} is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks. However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. \textbf{The extrapolation results of the velocity and the temperature differ from the direct measurements of the other studies. This suggests that below about \SI{20}{\Rs} PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur} (see Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the \textbf{solar wind is believed to be accelerated up to the height of the Alfvénic critical surface, which is predicted to lie on average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000}, scaling with solar activity within a range of between \SI{15}{\Rs} at solar minimum and \SI{30}{\Rs} at solar} maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

>>>>

	\caption{Fit coefficients \textbf{for the distance depending single lognormal function, based on equation (\ref{eq:single_lognormal_fit_function}) combined with (\ref{eq:power_function})}, respectively double lognormal \textbf{function (\ref{eq:double_lognormal_fit_function})} for the velocity from the combined Helios data. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value.} The seasonal variations are calculated from Earth's orbital solar distance variation and the derived exponents.}

@@

\textbf{The power law lognormal models and the power law double lognormal model for the velocity, which result from the fitting, are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot} together with their median values.} The model’s magnetic field strength is broader around values of \SI{40}{nT} at the lower distance boundary than the data's frequency distribution implies. This behavior is expected because of the distance independent shape approximation applied. The velocity and temperature models’ upper values generally show a higher abundance than the actual data, see also zoom boxes in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}. \textbf{The increasing high velocity tail with distance arises} from using the same exponent for both slow and fast components. This effect is not seen in the data, more specifically, not only the slowest wind but also the fastest wind is expected to converge to more average speeds \citep{Sanchez-Diaz2016}.
%\citep{Sanchez-Diaz2016} p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction)

@@

%\subsection{Distance scaling law variations}
%\label{sec:distance_scaling_law_variations}
\textbf{This solar wind model} for the radial distance dependence is representative for the time of the Helios observations around the rise of solar cycle~21. The \textbf{variations of the} yearly power law fit exponents \textbf{from fitting the solar distance dependency (\ref{eq:power_function})} are shown in Fig.~\ref{fig:yearly_gradients_c} together with the yearly SSN for the time period \numrange{1974}{1982}. It can be seen that during the Helios time period there might be some systematic  variation of the exponents with solar activity---at least for the velocity and temperature exponents.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/yearly_gradients_c.pdf}}
	\caption{Helios yearly variation of the solar wind \textbf{parameter power exponents for the dependence on radial distance} together with the SIDC 13-month smoothed monthly SSN. The weighted standard deviations and average values for all years are indicated by the shaded areas. In this plot the 21~days since Helios launch in the year 1974 are omitted because a distance range of merely \SIrange{0.95}{0.98}{\au} was covered that year.}
	\label{fig:yearly_gradients_c}
\end{figure}
%yearly variation of the solar wind parameters' average fit function exponents
However, for simplicity we assume, that the distance scaling laws can be treated as time independent and include the calculated exponents’ yearly variations $\Delta e$, summarized in Table~\ref{tab:mean_median_fit_parameter}, as relative uncertainties.

Since we neglect possible variations of the distance scaling laws, we combine the frequency distribution’s median solar activity dependency (\ref{eq:median_with_ssn}) derived for \SI{1}{\au} from the OMNI data with the power law exponents (\ref{eq:power_function}) derived from the Helios data:
\begin{align}
	x_\text{med}(ssn,r) &= (a_\text{med} \cdot ssn + b_\text{med}) \cdot r^{e'}	\,.	\label{eq:general_sw_model}
\end{align}
Thus\textbf{, implementing the median and average relations into the lognormal function (\ref{eq:single_lognormal_fit_function}),} we obtain the combined model function $W'''(x,ssn,r)$ and for the velocity $W_\text{II}'''(x,ssn,r)$ with the double lognormal function (\ref{eq:double_lognormal_fit_function}). \textbf{The corresponding median and average relations for each solar wind parameter, based on values resulting from our analyses, are listed below. Their numerical values are the fit parameters from Table~\ref{tab:ssn_fit_parameters} and the exponents from Table~\ref{tab:extrapolation_model_fit_parameters}.}

\textbf{\begin{itemize}
	\item The magnetic field strength relations, depending on solar activity and solar distance, are:
	\begin{align}
		B_\text{med}(ssn,r) &= (\SI{0.0131}{\nT} \cdot ssn + \SI{4.29}{\nT}) \cdot r^{-1.66}	\,,	\label{eq:median_B}\\
		B_\text{avg}(ssn,r) &= 1.0879 \cdot B_\text{med}(ssn,r)\,.
	\end{align}
	\item The proton velocity relations for the slow and fast components, depending on solar distance, are:
	\begin{align}
		v_\text{med}^\text{slow}(r) &= \SI{363}{\km\per\s} \cdot r^{0.099}	\,,&	&	&v_\text{med}^\text{fast}(r) &= \SI{483}{\km\per\s} \cdot r^{0.099}	\,,	\label{eq:median_v}\\
		v_\text{avg}^\text{slow}(r) &= 1.0101 \cdot v_\text{med}^\text{slow}(r)	\,,&	&	&v_\text{avg}^\text{fast}(r) &= 1.023 \cdot v_\text{med}^\text{fast}(r)	\,.
	\end{align}
	The share of both components is found to be balanced with solar activity:
	\begin{align}
		c(ssn) = -0.00180 \cdot ssn + 0.64	\,.	\label{eq:balance_v}
	\end{align}
	\item The derived relations of the proton density are:
	\begin{align}
		n_\text{med}(ssn,r) &= \left(\SI{0.0038}{\per\cm\cubed} \cdot ssn + \SI{4.50}{\per\cm\cubed}\right) \cdot r^{-2.11}	\,,	\label{eq:median_n}\\
		n_\text{avg}(ssn,r) &= 1.305 \cdot n_\text{med}(ssn,r)\,.
	\end{align}
	\item The derived proton temperature relations are:
	\begin{align}
		T_\text{med}(ssn,r) &= (\SI{197}{\K} \cdot ssn + \SI{57300}{\K}) \cdot r^{-1.10}	\,,	\label{eq:median_T}\\
		T_\text{avg}(ssn,r) &= 1.654 \cdot T_\text{med}(ssn,r)\,.
	\end{align}
\end{itemize}}

\textbf{These relations average} over seasonal variations because \textbf{they are} based on yearly data. As the OMNI data are time-shifted to the nose of the Earth’s bow shock, this leads to yearly solar distance variations of \SI{+-1.67}{\%} as it orbits the Sun. The resulting maximal solar wind parameter variation amplitudes over the year can thus be derived from the derived power law exponents. They are estimated to be smaller than \SI{4}{\percent} as seen in Tab.~\ref{tab:extrapolation_model_fit_parameters}. \citet{Bruno1986} \textbf{and} \citet{Balogh1999} have pointed out, that the solar wind parameters vary with latitudinal separation from the heliospheric current sheet (HCS). Its position in heliographic latitude is highly variable around the solar equator \citep{Schwenn1990}, furthermore, the Earth’s orbit varies over the course of the year by \SI{+-7.2}{\degree} in latitude. \textbf{Since this latitudinal separation is highly variable, and requires significant effort to calculate for an extended time series, we have ignored this aspect in this analysis.}	%Balogh et al. (p.~162, 1999);	 (Schwenn 1990, p.~126)

@@

	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. \textbf{The models obtained from Helios and OMNI measurements} are extrapolated to the PSP region \textbf{for SSN values between} solar minimum and maximum, i.e., $0 \le ssn \le 200$\textbf{. The lower edges of the shaded areas correspond to solar minimum, the upper edges to solar maximum.} Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \textbf{\citet{Sheeley1997} and \citet{Wang2000}}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}

@@

The magnetic field strength is found to increase from median values of about \SI{43}{\nT} at \SI{0.25}{\au} to \SI{715}{\nT} at \SI{0.046}{\au} for a SSN of 0. Taking a SSN of 200 increases the values to \SI{69}{\nT} and \SI{1152}{\nT}. Our extrapolation results are slightly flatter than those derived from the analytical magnetic field model by \citet{Banaszkiewicz1998}, who constructed an analytic dipole plus quadrupole plus current sheet (DQCS) model for solar minimum. \textbf{It should be noted that one cannot straightforeward compare the absolute values of our study with the values obtained by \citet{Banaszkiewicz1998} because the DQCS model assumes solar wind originating from coronal holes at higher heliographic latitudes only, neglecting the slow solar wind belt.}
%Since we cannot compare with their absolute values, because their DQCS model assumes fields of \SI{3.1}{\nT} at \SI{1}{\au} assumed to originate from coronal holes with boundaries at \SI{+-60}{\degree} latitude, excluding the fields from the equatorial slow wind belt.

@@

The proton density increases from about \SI{84}{\per\cm\cubed} at \SI{0.25}{\au} to about \SI{3018}{\per\cm\cubed} at \SI{0.046}{\au} \textbf{6~years after a SSN of 0 occurred}. Being almost independent of the SSN the values for a SSN of 200 are only \SI{17}{\%} larger. The results are in good agreement with those of \citet{Leblanc1998}, who derived an electron density model from type~III radio burst observations. Their model shows that the density distance dependency scales with $r^{-2}$ and steepens just below \SI{10}{\Rs} with $r^{-6}$. \textbf{We assumed a solar wind helium abundance of \SI{5}{\%} to convert these electron densities to proton densities.}

@@

%\subsection{PSP solar wind environment estimation}
Implementing the \textbf{SSN predictions} for the PSP mission time and the orbital trajectory data, \textbf{we can infer} which solar wind parameter magnitudes can be expected. Figs.~\ref{fig:SPP_perihelia_prediction_f_plot} and \ref{fig:SPP_perihelia_prediction_nearest_f_plot} show \textbf{the median values (\ref{eq:median_with_ssn}) of} the considered different solar wind parameters for 12-day periods, comprising the first perihelion in Novemver 2018 and the \textbf{first} closest perihelion in December 2024.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) during 12~days in 2018 with PSP's first perihelion at about \SI{0.16}{\au}. For the velocity the combined median is calculated and also the SSN independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_f_plot}
\end{figure}
% SSN sources and error ranges for first perihelion
% magnetic field:
% SWPC prediction - 0 years
% SWPC expected SSN ranges
% velocity:
% actual monthly SSN - 3 years
% not plotted
% density:
% actual monthly SSN - 6 years
% SIDC given std. dev. from stations
% temperature:
% actual monthly SSN - 3 years
% SIDC given std. dev. from stations
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) \textbf{during} 12~days in 2024 with PSP's \textbf{22nd (the first closest)} perihelion at \SI{0.0459}{\au}. For the velocity the combined median is calculated and also the SSN independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_nearest_f_plot}
\end{figure}
% SSN sources and error ranges for first closest perihelion
% magnetic field:
% actual monthly SSN - 11 years
% half/twice SSN
% velocity:
% actual monthly SSN - 14 years
% not plotted
% density:
% SWPC prediction - 0 years
% SWPC expected SSN ranges
% temperature:
% actual monthly SSN - 14 years
% half/twice SSN
In the beginning of the mission median values of about \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K} are estimated to be measured at \SI{0.16}{\au}, increasing to about \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K} during the \textbf{first} closest approach at \SI{0.046}{\au}. \textbf{Monthly SSNs---shifted by the time lags specific to the solar wind parameters---are used in the calculation of the solar wind predictions. These SSNs are either actual smoothed values from the SIDC with their reported standard deviations, short-term predictions from the SWPC with their expected ranges or actual smoothed values from the SIDC shifted by 11~years with half/twice their values as uncertainties. The error bands given in both figures, calculated from error propagation, include these SSN ranges and the derived fit parameter errors.}
%B
% 87.3642361846083
%N
% 213.613795298952	first
% 2950.67837375615	nearest
% 2925.63248639046	SSN 0
% 3421.59065204775	SSN 200
%T
% 503136.759398494

\textbf{Finally the estimated solar wind environment can be derived from the function $W'''(x,ssn,r)$. The estimated frequency distributions of the four solar wind parameters at PSP's first and 22nd (first closest) perihelia are plotted in Fig.~\ref{fig:SPP_sw_distributions}.}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_sw_distributions.pdf}}
	\caption{\textbf{Frequency distributions of the four solar wind parameters (same as in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}) and those estimated with the solar wind model for PSP's first and 22nd (first closest) perihelion. In these figures the frequencies of both extrapolated curves are scaled for visibility to the same height as the \SI{1}{\au} distribution.}}
	\label{fig:SPP_sw_distributions}
\end{figure}
\textbf{Note again that the velocity and temperature distributions for the 22nd perihelion are only upper limits and the actual by PSP to be encountered values are expected to be smaller.}

\section{Discussion and summary}
\label{sec:discussion_and_summary}
\textbf{The scientific objective of this study, being part of the German contribution Coronagraphic German and US Solar Probe Plus Survey (CGAUSS) to the Wide field Imager for Solar Probe (WISPR), is to model the solar wind environment for the Parker Solar Probe mission (formerly Solar probe Plus) to be launched mid 2018.}

@@

\textbf{For determining solar activity and solar distance dependent relations for the median and average solar wind values, we could have used the simpler approach of combining the radial dependence of averaged Helios data with averaged \SI{1}{\au} OMNI data scaled with the SSN. It is expected that the results of a simpler analysis would have similar distance scaling results, as can be inferred from the exponents in Tables~\ref{tab:mean_median_fit_parameter} and \ref{tab:extrapolation_model_fit_parameters}. However, in our study we are not only interested in averages but rather in bulk distributions, that is, the whole range of values that might occur. For the determination of the frequency distributions the use of the more complex fit model is important, because the distance between median and average values determines the width of the lognormal distributions. With our approach we are able to control and fit this distance.}

\textbf{It} is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks. However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. \textbf{The extrapolation results of the velocity and the temperature differ from the direct measurements seen in existing studies. This suggests that below about \SI{20}{\Rs} PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur} (see Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the \textbf{solar wind is assumed to be accelerated up to the height of the Alfvénic critical surface, which is predicted to lie on average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000}, scaling with solar activity within a range of between \SI{15}{\Rs} at solar minimum and \SI{30}{\Rs} at solar} maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).
